---
title: "demulticoder rhododendron rhizobiome workflow"
output: html_document
date: "2024-06-18"
---
## Demonstration on how to use demulticoder on a dataset that is actually three separate datasets (RPS10, ITS, and 16S)

### Load package
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/benchmark_demulticoder/demulticoder")
devtools::install_github("grunwaldlab/demulticoder", force=TRUE)
devtools::load_all("~/demulticoder")
library("demulticoder")
set.seed(1)
```

### Step 1-Remove N's and create directory structure for downstream steps
```{r prepare reads}
outputs<-prepare_reads(
  data_directory = "data", 
  output_directory = "outputs", 
  tempdir_path="temp_outputs",
  tempdir_id = "temp_files",
  overwrite_existing = FALSE)
```

### Step 2-Run Cutadapt to remove primers and then trim reads with DADA2 filterAndTrim function 
```{r Remove primers and trim reads}
cut_trim(
  outputs,
  cutadapt_path="/usr/bin/cutadapt",
  overwrite_existing = FALSE)
```

### Step 3-Core ASV inference step
```{r ASV inference}
make_asv_abund_matrix(
  outputs,
  overwrite_existing = FALSE)
```

## Step 4-Assign taxonomy step
```{r assign taxonomy step}
  assign_tax(
    outputs,
    asv_abund_matrix,
    db_rps10 = "release_1.fa",
    db_its = "sh_general_release_dynamic_18.07.2023.fasta",
    retrieve_files=TRUE,
    overwrite_existing=TRUE)
```

## Step 5-convert asv matrix to taxmap and phyloseq objects with one function
```{r convert matrix to other formats}
objs<-demulticoder::convert_asv_matrix_to_objs(outputs, save_outputs=TRUE, overwrite_existing = TRUE)
```

### Let's collect info on R configuration and associated package versions that are downloaded
```{r}
sessioninfo::session_info()
```