---
title: "Pooled amplicon example"
output: 
  rmarkdown::html_vignette:
    fig_path: "man/figures"
vignette: >
  %\VignetteIndexEntry{Pooled amplicon example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  fig.path = "./"
)
```

```{r setup}
library(phyloseq); packageVersion("phyloseq")
library(ggplot2); packageVersion("ggplot2")
library(readr); packageVersion("readr")
library(purrr); packageVersion("purrr")
library(furrr); packageVersion("furrr")
library(dplyr); packageVersion("dplyr")
library(stringr); packageVersion("stringr")
library(metacoder); packageVersion("metacoder")
library(data.table); packageVersion("data.table")
library(decontam); packageVersion("decontam")
library(tidyr); packageVersion("tidyr")
library(purrr); packageVersion("purrr")
library(forcats); packageVersion("forcats")
library(Biostrings); packageVersion("Biostrings")
library(magick); packageVersion("magick")
library(pdftools);packageVersion("pdftools")
knitr::opts_knit$set(root.dir = "~/benchmark_demulticoder")
```

### Demonstration of how to use demulticoder on a dataset that has pooled amplicons (RPS10 and ITS), and how generalized analysis would be done if you didn't use demulticoder. A more nuanced analysis is shown in 'pooled_amplicoin_analysis_demulticoder.Rmd'

### Let's also load files from the standard worfklows for its and dataset rps10 datasets, and combine. We'll then make a combined matrix that is analogous to what was done with the demulticoder analysis
```{r load files standard workflow}
taxa_its<- read.table("standard_workflow/its/data/its_taxa.out", sep="\t", stringsAsFactors=F,header=T)

taxa_filt_its <- taxa_its %>%
  mutate(across(starts_with("tax."), ~ ifelse(get(str_replace(cur_column(), "tax.", "boot.")) < 0, NA, .))) %>%
  set_names(str_replace(names(.), "tax.", ""))

remove_prefix_its <- function(data) {
  data <- gsub("[a-z]__", "", data)  
  return(data)
}

taxa_filt_its[] <- lapply(taxa_filt_its, remove_prefix_its)

revise_species_its <- function(genus, species) {
  revise_species_its <- ifelse(is.na(species), NA, paste(genus, species, sep = "_"))
  return(revise_species_its)
}

taxa_filt_its$Species <- revise_species_its(taxa_filt_its$Genus, taxa_filt_its$Species)
taxa_its_df <- as.data.frame(taxa_filt_its)
taxa_its_df$sequence <- rownames(taxa_its_df)
taxa_its_df <- taxa_its_df[, c("sequence", setdiff(names(taxa_its_df), "sequence"))]
rownames(taxa_its_df) <- NULL

#Load ASV matrix now
seqtab_nochim_its <- read.table("standard_workflow/its/data/its_seqtab_nochim.out")
seqtab_nochim_its <- t(seqtab_nochim_its)
seqtab_nochim_its_df <- as.data.frame(seqtab_nochim_its)
seqtab_nochim_its_df$sequence <- rownames(seqtab_nochim_its_df)
seqtab_nochim_its_df <- seqtab_nochim_its_df[, c("sequence", setdiff(names(seqtab_nochim_its_df), "sequence"))]
rownames(seqtab_nochim_its_df) <- NULL


#Combine matrices
combined_df_its <- merge(seqtab_nochim_its_df, taxa_its_df, by = "sequence", all = TRUE)

taxonomic_cols <- c("Kingdom", "Phylum", "Class", "Order", 
                    "Family", "Genus", "Species")

bootstrap_cols <- c("boot.Kingdom", "boot.Phylum", "boot.Class", "boot.Order", 
                    "boot.Family", "boot.Genus", "boot.Species")

dada2_tax <- mapply(function(tax, boot, level) {
  paste(tax, boot, level, sep = "--")
}, combined_df_its[, taxonomic_cols], combined_df_its[, bootstrap_cols], 
   c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species"))

combined_df_its$dada2_tax <- apply(dada2_tax, 1, paste, collapse = ";")

#Remove all other tax and boot columns
combined_df_its <- combined_df_its[, !(names(combined_df_its) %in% c(taxonomic_cols, bootstrap_cols))]

# Reorder columns to put 'taxa_bootstrap_combined' after 'sequence'
combined_df_its <- combined_df_its[, c("sequence", "dada2_tax", setdiff(names(combined_df_its), c("sequence", "dada2_tax")))]

#Add domain-100-Eukaryota; prefix before Kingdom
combined_df_its$dada2_tax <- gsub("Fungi", "Eukaryota--100--Domain;Fungi", combined_df_its$dada2_tax)

#Load metadata
samdf_its<- read.csv("standard_workflow/its/data/metadata_its.csv")

#Save file
write.csv(combined_df_its, "standard_workflow/its/data/its_abundance_matrix_combined_dada2workflow.csv")
```

### Track number of reads throughout the pipeline
```{r track reads through DADA2 analysis, include=TRUE}
track_reads_standard_wf_its <- read.table("standard_workflow/its/data/its_trackreads.out", sep="\t", stringsAsFactors = F, header=T)
```

### Let's also do the same for the rps10 dataset
```{r load files standard workflow}
taxa_rps10<- read.table("standard_workflow/rps10/data/rps10_taxa.out", sep="\t", stringsAsFactors=F,header=T)

taxa_filt_rps10 <- taxa_rps10 %>%
  mutate(across(starts_with("tax."), ~ ifelse(get(str_replace(cur_column(), "tax.", "boot.")) < 0, NA, .))) %>%
  set_names(str_replace(names(.), "tax.", ""))

remove_prefix_rps10 <- function(data) {
  data <- gsub("[a-z]__", "", data)  
  return(data)
}

taxa_filt_rps10[] <- lapply(taxa_filt_rps10, remove_prefix_rps10)

taxa_rps10_df <- as.data.frame(taxa_filt_rps10)
taxa_rps10_df$sequence <- rownames(taxa_rps10_df)
taxa_rps10_df <- taxa_rps10_df[, c("sequence", setdiff(names(taxa_rps10_df), "sequence"))]
rownames(taxa_rps10_df) <- NULL

#Load ASV matrix now
seqtab_nochim_rps10 <- read.table("standard_workflow/rps10/data/rps10_seqtab_nochim.out")
rownames(seqtab_nochim_rps10) <- gsub("__", "_", rownames(seqtab_nochim_rps10))

seqtab_nochim_rps10 <- t(seqtab_nochim_rps10)
seqtab_nochim_rps10_df <- as.data.frame(seqtab_nochim_rps10)
seqtab_nochim_rps10_df$sequence <- rownames(seqtab_nochim_rps10_df)
seqtab_nochim_rps10_df <- seqtab_nochim_rps10_df[, c("sequence", setdiff(names(seqtab_nochim_rps10_df), "sequence"))]
rownames(seqtab_nochim_rps10_df) <- NULL

#Combine matrices
combined_df_rps10 <- merge(seqtab_nochim_rps10_df, taxa_rps10_df, by = "sequence", all = TRUE)

taxonomic_cols <- c("Kingdom", "Phylum", "Class", "Order", 
                    "Family", "Genus", "Species")

bootstrap_cols <- c("boot.Kingdom", "boot.Phylum", "boot.Class", "boot.Order", 
                    "boot.Family", "boot.Genus", "boot.Species")

dada2_tax <- mapply(function(tax, boot, level) {
  paste(tax, boot, level, sep = "--")
}, combined_df_rps10[, taxonomic_cols], combined_df_rps10[, bootstrap_cols], 
   c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species"))

combined_df_rps10$dada2_tax <- apply(dada2_tax, 1, paste, collapse = ";")

#Remove all other tax and boot columns
combined_df_rps10 <- combined_df_rps10[, !(names(combined_df_rps10) %in% c(taxonomic_cols, bootstrap_cols))]

# Reorder columns to put 'taxa_bootstrap_combined' after 'sequence'
combined_df_rps10 <- combined_df_rps10[, c("sequence", "dada2_tax", setdiff(names(combined_df_rps10), c("sequence", "dada2_tax")))]

#Add domain-100-Eukaryota; prefix before Kingdom
combined_df_rps10$dada2_tax <- gsub("Stramenopila", "Eukaryota--100--Domain;Stramenopila", combined_df_rps10$dada2_tax)

#Load metadata
samdf_rps10<- read.csv("standard_workflow/rps10/data/metadata_rps10.csv")

#Save file
write.csv(combined_df_rps10, "standard_workflow/rps10/data/rps10_abundance_matrix_combined_dada2workflow.csv")

#Load DADA2 metadata file
samdf_rps10<- read.csv("standard_workflow/rps10/data/metadata_rps10.csv")
```

### Track number of reads throughout the pipeline
```{r track reads through DADA2 analysis, include=TRUE}
track_reads_standard_wf_rps10 <- read.table("standard_workflow/rps10/data/rps10_trackreads.out", sep="\t", stringsAsFactors = F, header=T)
```

### Finally Let's first combine rps10 and its data into fully combined matrices
```{r combine dataframes}
sample_cols_its <- setdiff(names(combined_df_its), c("sequence", "dada2_tax"))
sample_cols_rps10 <- setdiff(names(combined_df_rps10), c("sequence", "dada2_tax"))

if(!all(sample_cols_its == sample_cols_rps10)) {
  stop("Sample columns do not match between ITS and RPS10 dataframes!")
}

combined_final <- rbind(
  combined_df_its[, c("sequence", "dada2_tax", sample_cols_its)],  # ITS data
  combined_df_rps10[, c("sequence", "dada2_tax", sample_cols_rps10)]  # RPS10 data
)
`
write.csv(combined_final, "final_combined_abundance_matrix_standard_workflow.csv")
```

### Now, let's proceed with our analogous analysis to demulticoder analysis

