---
title: "Pooled amplicon analysis-directcompare"
output:
  html_document: default
  pdf_document: default
date: "2024-12-24"
---

```{r setup}
library(phyloseq); packageVersion("phyloseq")
library(ggplot2); packageVersion("ggplot2")
library(readr); packageVersion("readr")
library(tidyr); packageVersion("tidyr")
library(purrr); packageVersion("purrr")
library(furrr); packageVersion("furrr")
library(dplyr); packageVersion("dplyr")
library(stringr); packageVersion("stringr")
library(forcats); packageVersion("forcats")
library(metacoder); packageVersion("metacoder")
library(data.table); packageVersion("data.table")
library(decontam); packageVersion("decontam")
library(Biostrings); packageVersion("Biostrings")
library(magick); packageVersion("magick")
library(vegan); packageVersion("vegan")
library(pdftools);packageVersion("pdftools")
library(vegan); packageVersion("vegan")
library(grid); packageVersion("grid")
library(cowplot); packageVersion("cowplot")
library(gridExtra); packageVersion("gridExtra")
knitr::opts_knit$set(root.dir = "~/benchmark_demulticoder")
```

### Alpha diversity-Let's do a direct comparison of mean inverse Simpson diversity between workflows
```{r alpha diversity comp}
stats_diversity_demulticoder<-read.csv("demulticoder/results/alpha_diversity.csv")
stats_diversity_sf <- read.csv("standard_workflow/combined_results/alpha_diversity.csv")

stats_diversity_demulticoder <- stats_diversity_demulticoder[order(stats_diversity_demulticoder$sample_name), ]
stats_diversity_sf <- stats_diversity_sf[order(stats_diversity_sf$sample_name), ]

stats_diversity_sf <- stats_diversity_sf[match(stats_diversity_demulticoder$sample_name, stats_diversity_sf$sample_name), ]

if (!all(stats_diversity_demulticoder$sample_name == stats_diversity_sf$sample_name)) {
  stop("Sample names do not match between the two data frames.")
}

sample_name <- stats_diversity_sf$sample_name
#raw_demulticoder <- stats_diversity_demulticoder$raw_count
#raw_sf <- stats_diversity_sf$raw_count
#raw_count_diff <- stats_diversity_demulticoder$raw_count - stats_diversity_sf$raw_count
richness_sf <- stats_diversity_sf$richness
richness_demulticoder <- stats_diversity_demulticoder$richness
richness_diff <- abs(stats_diversity_sf$richness - stats_diversity_demulticoder$richness) 
shannon_demulticoder <- stats_diversity_demulticoder$shannon
shannon_sf <- stats_diversity_sf$shannon
shannon_diff <- abs(stats_diversity_sf$shannon - stats_diversity_demulticoder$shanno)  
invsimpson_demulticoder <- stats_diversity_demulticoder$invsimpson
invsimpson_sf <- stats_diversity_sf$invsimpson
invsimpson_diff <- abs(stats_diversity_sf$invsimpson - stats_diversity_demulticoder$invsimpson)

difference_df <- data.frame(sample_name, richness_sf, richness_demulticoder,  richness_diff, shannon_sf, shannon_demulticoder, shannon_diff, invsimpson_sf, invsimpson_demulticoder, invsimpson_diff)

# save csv
write.csv(difference_df, "demulticoder_comp_sf/results/alpha_diversity_comparison.csv", row.names = FALSE)

# Let's do a paired t-test and wilcoxon test for each of the alpha diversity metrics
# Raw count
wilcox.test(stats_diversity_demulticoder$raw_count, stats_diversity_sf$raw_count,paired = TRUE)
t.test(stats_diversity_demulticoder$raw_count, stats_diversity_sf$raw_count,paired = TRUE)

# Richness
wilcox.test(stats_diversity_demulticoder$richness, stats_diversity_sf$richness,paired = TRUE)
t.test(stats_diversity_demulticoder$richness, stats_diversity_sf$richness,paired = TRUE)

# Shanonn index
wilcox.test(stats_diversity_demulticoder$shannon, stats_diversity_sf$shannon,paired = TRUE)
t.test(stats_diversity_demulticoder$shannon, stats_diversity_sf$shannon,paired = TRUE)

#Inverse simpson
wilcox.test(stats_diversity_demulticoder$invsimpson, stats_diversity_sf$invsimpson,paired = TRUE)
t.test(stats_diversity_demulticoder$invsimpson, stats_diversity_sf$invsimpson,paired = TRUE)
```

### Make similar table for nmds data
```{r beta diversity comp}
# Read in the demulticoder data
nmds_demulticoder<-read.csv("demulticoder/results/compiled_stats.csv")
nmds_sf <- read.csv("standard_workflow/combined_results/compiled_stats.csv")

nmds_demulticoder <- nmds_demulticoder[order(nmds_demulticoder$sample_name), ]
nmds_sf <-nmds_sf[order(nmds_sf$sample_name), ]

nmds_sf <- nmds_sf[match(nmds_demulticoder$sample_name, nmds_sf$sample_name), ]

if (!all(nmds_demulticoder$sample_name == nmds_sf$sample_name)) {
  stop("Sample names do not match between the two data frames.")
}

nmds1_diff <- abs(nmds_sf$NMDS1 - nmds_demulticoder$NMDS1)
nmds2_diff <- abs(nmds_sf$NMDS2 - nmds_demulticoder$NMDS2)
sample_name <- nmds_demulticoder$sample_name
nmds1_demulticoder <- nmds_demulticoder$NMDS1
nmds1_sf <- nmds_sf$NMDS1
nmds2_demulticoder <- nmds_demulticoder$NMDS2
nmds2_sf <- nmds_sf$NMDS2
expected_innoc<-nmds_demulticoder$only_expected_innoc

#Add to new table
nmds_df <- data.frame(sample_name, expected_innoc, nmds1_sf, nmds1_demulticoder, nmds1_diff, nmds2_sf, nmds2_demulticoder,nmds2_diff)

# save csv
write.csv(nmds_df, "demulticoder_comp_sf/results/beta_diversity_comparison.csv", row.names = FALSE)
```


### Make facetted figure with alpha diversity and nmds data loaded above
Start with making demulticoder boxplots-alpha diversity
```{r facetted figure-alpha div demulticoder}
plotted_factors <- c('Organism' = 'organism', 'Flooded' = 'flooded', 'Pathogen Concentration' = 'path_conc', 'Trial' = 'experiment')

alpha_plot_data <- plotted_factors %>%
  map2_dfr(names(plotted_factors), function(factor_col, factor_name) {
    out <- stats_diversity_demulticoder
    out$factor <- factor_name
    out$value <- as.character(out[[factor_col]])
    return(out)
  }) %>%
  mutate(path_conc = factor(path_conc, levels = sort(unique(path_conc)), labels = paste(sort(unique(path_conc)), 'CFU/g'), ordered = TRUE)) %>%
  filter(sample_type == 'Sample') %>%
  select(sample_name, factor, value, invsimpson) %>%
  tidyr::gather(key = "index", value = "diversity", -sample_name, -factor, -value) %>%
  mutate(value = forcats::fct_relevel(ordered(value), "Control", "Cin", "Plu", "Cry"))

alpha_plot_data <- alpha_plot_data %>%
  mutate(value = recode(value, "FALSE" = "Not Flooded", "TRUE" = "Flooded", "Control" = "Con"))

anova_and_hsd <- function(x) {
  anova_result <- aov(diversity ~ value, x)
  tukey_result <- agricolae::HSD.test(anova_result, "value", group = TRUE)
  group_data <- tukey_result$groups[order(rownames(tukey_result$groups)),]
  group_key <- setNames(group_data$groups, rownames(group_data))
  group_key[as.character(x$value)]
}
alpha_plot_data$group <- unlist(map(split(alpha_plot_data, alpha_plot_data$factor)[unique(alpha_plot_data$factor)], anova_and_hsd))

alpha_subplot_demulticoder <- ggplot(alpha_plot_data, aes(x = value, y = diversity)) +
  geom_boxplot(color = "black", fill = "white") +
  geom_text(aes(x = value, y = max(diversity) + 2, label = group), col = 'black', size = 3) +
  facet_grid(~ factor, scales = "free") +
  labs(x = NULL, y = 'Diversity (Inverse Simpson)') +
  guides(color = "none") +
  theme(panel.background = element_rect(fill = "white"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_line(color = "gray"),
        axis.text.x = element_text(color = "black"),
        axis.text.y = element_text(color = "black"),
        axis.title.y = element_text(color = "black"),
        legend.position = "bottom",
        panel.border = element_rect(color = "black", fill = NA, size = 1),
        strip.background = element_rect(fill = "white", color = "white"),
        strip.text = element_text(color = "black"))

print(alpha_subplot_demulticoder)
```

### Start with making standard workflow boxplots-alpha diversity
```{r facetted figure-alpha div standard wf}
plotted_factors <- c('Organism' = 'organism', 'Flooded' = 'flooded', 'Pathogen Concentration' = 'path_conc', 'Trial' = 'experiment')

alpha_plot_data <- plotted_factors %>%
  map2_dfr(names(plotted_factors), function(factor_col, factor_name) {
    out <- stats_diversity_sf
    out$factor <- factor_name
    out$value <- as.character(out[[factor_col]])
    return(out)
  }) %>%
  mutate(path_conc = factor(path_conc, levels = sort(unique(path_conc)), labels = paste(sort(unique(path_conc)), 'CFU/g'), ordered = TRUE)) %>%
  filter(sample_type == 'Sample') %>%
  select(sample_name, factor, value, invsimpson) %>%
  tidyr::gather(key = "index", value = "diversity", -sample_name, -factor, -value) %>%
  mutate(value = forcats::fct_relevel(ordered(value), "Control", "Cin", "Plu", "Cry"))

alpha_plot_data <- alpha_plot_data %>%
  mutate(value = recode(value, "FALSE" = "Not Flooded", "TRUE" = "Flooded", "Control" = "Con"))

anova_and_hsd <- function(x) {
  anova_result <- aov(diversity ~ value, x)
  tukey_result <- agricolae::HSD.test(anova_result, "value", group = TRUE)
  group_data <- tukey_result$groups[order(rownames(tukey_result$groups)),]
  group_key <- setNames(group_data$groups, rownames(group_data))
  group_key[as.character(x$value)]
}
alpha_plot_data$group <- unlist(map(split(alpha_plot_data, alpha_plot_data$factor)[unique(alpha_plot_data$factor)], anova_and_hsd))

alpha_subplot_sf <- ggplot(alpha_plot_data, aes(x = value, y = diversity)) +
  geom_boxplot(color = "black", fill = "white") +
  geom_text(aes(x = value, y = max(diversity) + 2, label = group), col = 'black', size = 3) +
  facet_grid(~ factor, scales = "free") +
  labs(x = NULL, y = 'Diversity (Inverse Simpson)') +
  guides(color = "none") +
  theme(panel.background = element_rect(fill = "white"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_line(color = "gray"),
        axis.text.x = element_text(color = "black"),
        axis.text.y = element_text(color = "black"),
        axis.title.y = element_text(color = "black"),
        legend.position = "bottom",
        panel.border = element_rect(color = "black", fill = NA, size = 1),
        strip.background = element_rect(fill = "white", color = "white"),
        strip.text = element_text(color = "black"))

print(alpha_subplot_sf)
```

### Start with making demulticoder nmds plots
```{r facetted figure-beta div-demulticoder}
set.seed(10)
nmds_factors <- c(Flooded = 'flooded', Organism = 'organism', 'Pathogen CFU/g' = 'path_conc', 'Trial' = 'experiment')

nmds_data <- nmds_demulticoder

nmds_data <- nmds_data %>%
  mutate(flooded = as.character(flooded),
         flooded = recode(flooded, "TRUE" = "Flooded", "FALSE" = "Not Flooded"),
         organism = recode(organism, "Control" = "Con"))

make_one_plot <- function(factor, name) {
  nmds_data %>%
    mutate(factor = as.character(nmds_data[[factor]]),
           NMDS1 = scales::rescale(NMDS1),
           NMDS2 = scales::rescale(NMDS2)) %>%
    mutate(factor = fct_relevel(ordered(factor), "Con", "Cin", "Plu", "Cry")) %>%
    ggplot(aes_string(x = "NMDS1", y = "NMDS2", color = "factor", label = "sample_name")) +
    geom_point(size = 1.5) +  # Adjust point size
    coord_fixed() +
    viridis::scale_color_viridis(discrete = TRUE, end = .9) +
    labs(color = NULL, x = "NMDS1", y = "NMDS2") +
    theme(panel.background = element_rect(fill = "white"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.text = element_text(size = 8),
          axis.title = element_text(size = 10),
          axis.ticks = element_line(color = "black"),
          plot.margin = unit(rep(0.04, 4), "cm"),
          legend.position = "bottom",
          legend.text = element_text(size = 10),
          legend.key.height = unit(0.25, 'cm'),
          legend.key.width = unit(0.1, 'cm'),
          legend.spacing = unit(0.05, 'cm'),
          legend.margin = margin(t = -0.2, unit = 'cm'),
          panel.border = element_rect(color = "black", fill = NA, size = 1)) +
    guides(color = guide_legend(nrow = 2, byrow = TRUE)) +
 guides(color = guide_legend(nrow = 2, byrow = TRUE)) +
    scale_x_continuous(breaks=c(0, 0.5, 1), labels=c(0, 0.5, 1)) +  # Adjust x-axis labels
    scale_y_continuous(breaks=c(0, 0.5, 1), labels=c(0, 0.5, 1))    # Adjust y-axis labels
}

nmds_subplots <- map2(nmds_factors, names(nmds_factors), make_one_plot)
nmds_plot_demulticoder <- ggpubr::ggarrange(plotlist = c(list(ggplot() + theme_void()), nmds_subplots),
                       nrow = 1, widths = c(0.15, 1, 1, 1, 1))

nmds_plot_demulticoder
```

### Start with making standard workflow nmds plots
```{r facetted figure-beta div-sf}
set.seed(1)
nmds_factors <- c(Flooded = 'flooded', Organism = 'organism', 'Pathogen CFU/g' = 'path_conc', 'Trial' = 'experiment')

nmds_data <- nmds_sf

nmds_data <- nmds_data %>%
  mutate(flooded = as.character(flooded),
         flooded = recode(flooded, "TRUE" = "Flooded", "FALSE" = "Not Flooded"),
         organism = recode(organism, "Control" = "Con"))

make_one_plot <- function(factor, name) {
  nmds_data %>%
    mutate(factor = as.character(nmds_data[[factor]]),
           NMDS1 = scales::rescale(NMDS1),
           NMDS2 = scales::rescale(NMDS2)) %>%
    mutate(factor = fct_relevel(ordered(factor), "Con", "Cin", "Plu", "Cry")) %>%
    ggplot(aes_string(x = "NMDS1", y = "NMDS2", color = "factor", label = "sample_name")) +
    geom_point(size = 1.5) +  # Adjust point size
    coord_fixed() +
    viridis::scale_color_viridis(discrete = TRUE, end = .9) +
    labs(color = NULL, x = "NMDS1", y = "NMDS2") +
    theme(panel.background = element_rect(fill = "white"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.text = element_text(size = 8),
          axis.title = element_text(size = 10),
          axis.ticks = element_line(color = "black"),
          plot.margin = unit(rep(0.04, 4), "cm"),
          legend.position = "bottom",
          legend.text = element_text(size = 10),
          legend.key.height = unit(0.25, 'cm'),
          legend.key.width = unit(0.1, 'cm'),
          legend.spacing = unit(0.05, 'cm'),
          legend.margin = margin(t = -0.2, unit = 'cm'),
          panel.border = element_rect(color = "black", fill = NA, size = 1)) +
    guides(color = guide_legend(nrow = 2, byrow = TRUE)) +
    scale_x_continuous(breaks=c(0, 0.5, 1), labels=c(0, 0.5, 1)) +  # Adjust x-axis labels
    scale_y_continuous(breaks=c(0, 0.5, 1), labels=c(0, 0.5, 1))    # Adjust y-axis labels
}


nmds_subplots <- map2(nmds_factors, names(nmds_factors), make_one_plot)
nmds_plot_sf <- ggpubr::ggarrange(plotlist = c(list(ggplot() + theme_void()), nmds_subplots),
                       nrow = 1, widths = c(0.15, 1, 1, 1, 1))

nmds_plot_sf
```

### Now facet into A-D
```{r facetted figure-beta div}
combined_div_plot <- plot_grid(alpha_subplot_demulticoder, alpha_subplot_sf, nmds_plot_demulticoder, nmds_plot_sf, 
                               ncol = 2, labels = c('A.', 'B.', 'C.', 'D.'), 
                               rel_heights = c(0.5, 0.5, 1, 1))

combined_div_plot <- combined_div_plot + 
  theme(plot.margin = unit(c(0.25, 0.25, 0.25, 0.25), "cm"))

ggsave(combined_div_plot, path = "demulticoder_comp_sf/figures", filename = "diversity_demulticoder_sf.pdf", 
       width = 12.5, height = 4.75, bg = "#FFFFFF")

ggsave(combined_div_plot, path = "demulticoder_comp_sf/figures", filename = "diversity_demulticoder_sf.svg", 
       width = 12.5, height = 4.75, bg = "#FFFFFF")
```


### Let's collect info on R configuration and associated package versions that are downloaded
```{r session info}
sessioninfo::session_info()
```